Extended Data

### 2022-05-22 doc/manuscript/Dow_manuscript_Nature.Rmd
Extended Data Legends go in main manuscript file
Removed these from ED file, added in main file. Also, incorporated footnote of EDT2 into legend. And fixed up some typos, etc.

>>>>

\newpage
# Legends for Extended Data Table & Figures {-}

**Extended Data Table 1 | Dominant broadleaf deciduous species at the Smithsonian Conservation Biology Institute (SCBI) and Harvard Forest, along with sample sizes included in our final analysis.**

**Extended Data Table 2 | Summary of parameters describing the seasonality and temperature sensitivity of broadleaf deciduous tree woody growth and canopy phenology at the Smithsonian Conservation Biology Institute (SCBI) and Harvard Forest.** Temperature sensitivity refers to the sensitivity of a parameter to the mean maximum temperature (*T~max~*) during the critical temperature window (_CTW_, Fig. 1). Because leaf phenology measurements were derived from satellite imagery and included both ring- and diffuse-porous trees, the selection of a critical window was less straightforward. We used the _CTW_ of the porosity group containing the dominant canopy species at each site: *Liriodendron tulipifera* (diffuse porous) at SCBI and *Quercus rubra* (ring porous) at Harvard Forest.

**Extended Data Table 3 | Summary of tree-ring chronologies analyzed and number of significant (at significance level = 0.05) positive or negative correlations of ring width index to monthly *T~max~* in univariate and multivariate analyses.**

**Extended Data Figure 1 | Seasonal patterns of forest canopy greenness (top row) and stem growth of ring- and diffuse-porous trees, represented as both relative and cumulative fractions of total annual growth (middle and bottom rows, respectively), at the Smithsonian Conservation Biology Institute (SCBI) and Harvard Forest.** In the top row, canopy greenness is characterized using the two band Enhanced Vegetation Index (EVI2), with each line representing a year between 2000 and 2018. For stem growth, each line represents the average growth over one year, as modeled based on a five-parameter logistic growth model to dendrometer band data. Dashed lines represent modeled DBH change which fell outside of the median DOY where predicted starting DBH (a) and predicted ending DBH (b) were reached. Solid lines represent DBH change attributable to stem growth.

**Extended Data Figure 2 | Landscapes of relationships between the day of year on which 25% of annual growth is achieved (*DOY~25~*) and temperature in prior weeks for ring- and diffuse-porous trees at the Smithsonian Conservation Biology Institute (SCBI) and Harvard Forest.** Shown are matrices of $\beta$ coefficients from first-order linear regressions between mean maximum temperature (*T~max~*) and *DOY~25~*. Window Open and Window Close indicate number of weeks prior to *DOY~25~* (listed in `r EDT_phenology_parameters`). Yellow shading indicates neutral relationships, while orange or red shading indicates that *DOY~25~* advances with increased *T~max~* over the given time window (negative $\beta$). Black circles indicate the critical temperature window selected based on minimization of *∆AICc*, the difference in Akaike Information Criterion corrected for small sample size relative to a null model. Critical temperature windows are listed in `r EDT_phenology_parameters`.

**Extended Data Figure 3 | Response of stem growth timing and rates to mean maximum temperatures (*T~max~*) during the spring critial temperature window (_CTW_) for ring- and diffuse-porous species at the Smithsonian Conservation Biology Institute (SCBI) and Harvard Forest.** _CTW_ was defined as the period over which *T~max~* was most strongly correlated with the day of year on which 25% of annual growth was achieved (*DOY~25~*; `r EDT_phenology_parameters`, `r EDF_climwin`). Shown are relationships between mean *T~max~* over the _CTW_ and days of the year on which 25%, 50%, and 75% total stem growth were achieved (*DOY~25~*, *DOY~50~*, *DOY~75~*, respectively; first row); the length of the peak growing season (*L~pgs~*; second row); maximum growth rate (*g~max~*; third row); and total seasonal radial stem growth (*∆DBH*; fourth row). Posterior predictions of each variable that did not include zero are represented with solid lines, while those that do include zero use dotted lines. The 95% credible intervals are represented by bands centered on the posterior mean for each year. For both species groups at both sites, *DOY~25~*, *DOY~50~*, and *DOY~75~* all declined significantly with mean *T~max~* during their respective _CTW_. Dots represent growth parameter values for individual tree-year combinations, which were derived by fitting a five-parameter logistic growth model to dendrometer band data.

**Extended Data Figure 4 | Response of stem growth timing and rates to mean maximum temperatures (*T~max~*) for the month most closely corresponding to the spring critial temperature window (_CTW_) for ring- and diffuse-porous species at the Smithsonian Conservation Biology Institute (SCBI) and Harvard Forest.** _CTW_ was defined as the period over which *T~max~* was most strongly correlated with the day of year on which 25% of annual growth was achieved (*DOY~25~*; `r EDT_phenology_parameters`, `r EDF_climwin`), and the most closely corresponding month was determined as that with the greatest number of days within the _CTW_. Shown are relationships between monthly *T~max~* and days of the year on which 25%, 50%, and 75% total stem growth were achieved (*DOY~25~*, *DOY~50~*, *DOY~75~*, respectively; first row); the length of the peak growing season (*L~pgs~*; second row); maximum growth rate (*g~max~*; third row); and total seasonal radial stem growth (*∆DBH*; fourth row). Posterior predictions of each variable that did not include zero are represented with solid lines, while those that do include zero use dotted lines. The 95% credible intervals are represented by bands centered on the posterior mean for each year. For both species groups at both sites, *DOY~25~*, *DOY~50~*, and *DOY~75~* declined significantly with April *T~max~*, with the exception of *DOY~25~* for ring-porous species at SCBI. Dots represent growth parameter values for individual tree-year combinations, which were derived by fitting a five-parameter logistic growth model to dendrometer band data.

**Extended Data Figure 5 | Map of sampling locations of tree-ring chronologies analyzed in this study.** Sites are colored by the xylem porosity type of species sampled: ring poroous (RP), diffuse porous (DP), or both. Sampling details are provided in `r chron_table`. Base map source is ggplot2.

**Extended Data Figure 6 | Sensitivity of annual growth, as derived from tree-rings, to monthly mean minimum temperatures (*T~min~*), for `r n_chronologies` chronologies from `r n_cores_sites` sites across eastern North America** (`r EDF_map `). Colors indicate the bootstrapped correlation between monthly *T~min~* and a dimensionless ring width index (_RWI_) derived from the multiple trees that form each chronology and emphasizing interannual variability associated with climate. Chronologies are grouped by xylem porosity and ordered by mean April *T~min~*. Plots are annotated to highlight records from our two focal sites, the Smithsonian Conservation Biology Institute (SCBI) and Harvard Forest (HF) (`r EDT_species`). Species analyzed and numbers of significant correlations to *T~min~* are summarized in `r EDT_tree_ring_summary`, and chronology details are given in `r chron_table`.

**Extended Data Figure 7 | Sensitivity of annual growth, as derived from tree-rings, to monthly mean maximum temperatures (*T~max~*) of the current and past year, for `r n_chronologies` chronologies from `r n_cores_sites` sites across eastern North America** (`r EDF_map `). Colors indicate the bootstrapped correlation between monthly *T~max~* and a dimensionless ring width index (_RWI_) derived from the multiple trees that form each chronology and emphasizing interannual variability associated with climate. Chronologies are grouped by xylem porosity and ordered by mean April *T~max~*. Plots are annotated to highlight records from our two focal sites, the Smithsonian Conservation Biology Institute (SCBI) and Harvard Forest (HF) (`r EDT_species`). Figure presents the same results as main manuscript Fig. 3 but extends back to include the previous year. Species analyzed and numbers of significant correlations to *T~max~* are summarized in `r EDT_tree_ring_summary`, and chronology details are given in `r chron_table`.
